Illuminating a Dark Lens : A Type Ia Supernova \\ Magnified by Galaxy Cluster Abell 2744

### 2015-05-16 snTomas.tex
more tweaks to discussion section. Updated some figs

In the case of lensed SNe, we can also reverse the experimental setup:
instead of using strong-lensing clusters to study distant SNe, we can
use the SNe as tools for examining the lenses \citep{Riehm:2011}.  The
most valuable transients for testing and improving cluster lens models
would be {\it strongly lensed} SNe that are resolved into multiple
images \citep{Holz:2001,Oguri:2003}.  Transients that are lensed into
multiple images can also become cosmological tools, as the measurement
of time delays between the images can provide cosmographic information
to constrain the Hubble parameter \citep{Refsdal:1964} and other
cosmological parameters \citep{Linder:2011}.  We have recently
observed the first example of a multiply-imaged
SN \citep{Kelly:2015}, and expect to detect the reappearance of this
object (called ``SN Refsdal'') within 5 years, delivering a precise
test of lens model predictions \citep{Oguri:2015,Sharon:2015,Diego:2015}.
Although detections of such objects are currently very
unlikely \citep{Li:2012}, they will become much more common in the
next decade \citep{Coe:2009,Dobke:2009}, and may be developed into an
important new cosmological tool.

@@

One of the key values in observing standard candles behind
gravitational lenses is in addressing the problem of the
mass-sheet degeneracy \citep{Falco:1985,Bradac:2004}.  This
degeneracy arises because one can introduce into a lens model an
unassociated sheet of uniform mass in front of or behind the
lens, without disturbing the primary observable quantities.  For
example, take a lens model with a given surface mass density $\kappa$,
and then transform the surface mass density to
$\kappa^{\prime}=(1-\lambda)\kappa+\lambda$ for any arbitrary value
$\lambda$.  Both the $\kappa$ and $\kappa^{\prime}$ models will
produce exactly the same values for all positional and shear
constraints from strong and weak lensing.  When lensed background
sources are available across a wide range of redshifts (as is the case
for the Abell 2744 cluster discussed here), it should in principle be
possible to account for the simplest form of a single mass sheet
within that redshift range.  However, there are more complex versions
of positional constraint
degeneracies \citep{Schneider:2014,Liesenborgs:2012}.  Such
degeneracies do not extend to the {\it absolute} magnification of a
background source's flux and size.  Therefore, an absolute measurement
of magnification from a standard candle or a standard ruler
\citep{Sonnenfeld:2011}  can break these fundamental degeneracies
\citep{Holz:2001}.

@@

SN \tomas\ was discovered in Hubble Space Telescope (\HST) observations
with the Advanced Camera for Surveys (ACS) in the F606W and F814W
bands (V and i), collected on UT 2014 May 15 as part of the Hubble
Frontier Fields (HFF) survey (PI:J.Lotz,
HST-PID:13495).\footnote{\url{http://www.stsci.edu/hst/campaigns/frontier-fields}}
The HFF program is a 3-year director's discretionary initiative that
is collecting 140 orbits of HST imaging (roughly 340 ksec) on six
massive galaxy clusters, plus 6 accompanying parallel fields.  Each
field is observed in 3 optical bands (ACS F435W, F606W and F814W) and
4 infrared (IR) bands (WFC3-IR F105W, F125W, F140W, and F160W),
although the optical and IR imaging campaigns are separated by $\sim$6
months. Abell 2744 was the first cluster observed, with IR imaging
spanning 2013 October-November, and optical imaging from 2014
May-July.  A composite image of the HFF data showing the SN is
presented in Figure~\ref{fig:DiscoveryImage}, and the
locations of the cluster center, the SN, and the presumed host galaxy
are given in Table~\ref{tab:Positions}.  The SN detection was made in
difference images constructed using template imaging of Abell 2744
from HST+ACS observations taken in 2009 (PI:Dupke, HST-PID:11689).

@@

From the preceding analysis, we conclude that \tomas\ is a Type Ia SN
at $z\approx1.35$.  At this redshift, the absorption at $\sim$8700\AA\
corresponds to the blended \ion{Ca}{2} H\&K features.
This \ion{Ca}{2} absorption is commonly seen in Type Ia SN spectra
near maximum light, although it is also prominent in the spectra of
Type Ib and Ic core collapse SNe (\CCSNe).  The red color of
the \tomas\ SED is qualitatively consistent with a redshift of $z>1$
-- although this information was not used by {\tt SNID} for the
template matching.  As we will see in Section~\ref{sec:PhotometricClassification},
this spectral classification of SN \tomas\ is reinforced by the
photometric information, which also supports classification as a Type
Ia SN at $z\approx1.35$.

@@

To refine the redshift and phase constraints on \tomas, we next fit
the spectrum with a custom spectral matching program that employs a
library of Type Ia SN SEDs.  This library is similar to the Type Ia
spectral set used by SNID, but also includes more recent SNe with
well-observed spectral time series that extend to rest-frame UV
wavelengths (e.g. SN 2011fe and 2014J).  We first use an approach
similar to the SNID algorithm: warping the pseudo-continuum of each
template spectrum by dividing out a 3$^{\rm rd}$-order polynomial to
match the observed SED of SN \tomas.  This approach will find
templates that have similar abundances and photospheric
velocities. With the redshift fixed at $z=1.3457$ we find the best fit
is a spectrum from the normal Type Ia SN 2014J \citep{Foley:2014},
with a $\chi^2$ per degree of freedom $\nu$ of $\chi^2/\nu=121.8/89$
shown in the top panel of Figure~\ref{fig:SpecFit}.  Allowing the
redshift as an additional free parameter, we still find results that
are consistent with the SNID fits and $<2\sigma$ from the host galaxy
redshift: $z=1.31\pm0.02$ and a phase of $0\pm3$ rest-frame days. The
best-fitting spectral template in this case is the normal Type Ia SN
2012cg \citep{Amanullah:2015} at $z=1.31$ with $\chi^2/\nu=98.5/88$
(second panel of Figure~\ref{fig:SpecFit}).

@@

With the type and redshift securely defined as a normal Type Ia SN at
$z=1.3457$, we now turn to fitting the light curve with Type Ia
templates to measure the distance modulus  and then derive
the gravitational lensing magnification.  In this process, we would
like to avoid introducing systematic uncertainties inherent to any
assumed cosmological model \citep[e.g.][]{Nordin:2014}.  To that end,
we follow \citealt{Patel:2014} and define the magnification by
comparing the measured distance modulus of SN \tomas\ against an
average distance modulus derived from a ``control sample'' of unlensed
Type Ia SNe at similar redshift.  This allows us to make only the
minimal assumption that the redshift-distance relationship for Type Ia
SN is smooth and approximately linear over a small redshift span,
which should be true for any plausible cosmological model.

@@

The unlensed sample comprises 22 spectroscopically confirmed Type Ia
SNe in the range $1.14\leq z \leq1.55$ from three HST surveys: 11 from GOODS\footnote{GOODS: the Great Observatories Origins Deep Survey, PI:Giavalisco, HST-PID:9425,9583} \citep{Riess:2007}, 7 from SCP\footnote{SCP: the Supernova Cosmology Project, PI:Perlmutter} \citep{Suzuki:2012}, and 4 from CANDELS\footnote{CANDELS: the Cosmic Assembly Near-infrared Deep Extragalactic Legacy Survey, PI:Faber \&\ Ferguson} \citep{Rodney:2012,Rodney:2014}.  Using the SALT2 and MLCS2k2
fitters as described above, we get distance modulus measures for every
object in this control sample.  We then fit a linear relationship for
distance modulus vs. redshift, and derive a prediction for the
distance modulus of a normal Type Ia SN at the redshift of SN \tomas\
(Figure~\ref{fig:MagnificationMeasurement}).
This predicted value is given in
Table~\ref{tab:MeasuredMagnification} under the ``Control'' column.
The difference between the observed distance of SN \tomas\ and this
control sample value is attributed to the magnification from
gravitational lensing:

@@

\section{Discussion}
\label{sec:Discussion}

%\subsection{Comparison to Model Predictions}
%\label{sec:ComparisonToModelPredictions}

@@

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{FIG/snTomas_magnifications}
\caption{ 
Comparison of the observed lensing magnification to predictions from
lens models. The vertical blue line shows the constraints from
SN \tomas\ derived in Section~\ref{sec:DistanceAndMagnification} using the
MLCS2k2 fitter, with a shaded region marking the total
uncertainty. Markers with horizontal error bars show the median
magnification and 68\%\ confidence region from each of the 15 lensing
models.  Circles indicate models that use only strong-lensing
constraints, while diamonds denote those that also incorporate
weak-lensing measurements.  Models shown with open markers were
constructed using only data available before the start of the Frontier
Fields observations.  Filled markers indicate models that use
additional input constraints, including new multiply imaged systems
and redshifts.  Points with magenta outline indicate ``unblind''
models, which were generated after the magnification of the SN was
known. The eight models shown in black at the top are from
the ``parametric'' family, and the six in green at the bottom are
``free-form.''  The grey dashed line marks the unweighted mean for the
set of all models, at $\mu=2.5$.
\label{fig:LensingTest} }
\end{center}
\end{figure}

Table~\ref{tab:PredictedMagnifications} gives each model's predicted
magnification and uncertainty for a source at $z=1.3457$ and
at the position of \tomas.  These models represent a broad sampling of
the techniques and assumptions that can be applied to the modeling of
mass distributions in galaxy clusters.  The first eight models in this
table are so-called {\it parametric}\footnote{Although this
nomenclature is becoming standard in the literature, it is somewhat
misleading, as the pixels or grid cells in free-form models are
effectively parameters as well. Perhaps ``simply-parameterized'' would
be more accurate, though we adopt the more common usage here.}
models, and the bottom seven are {\it free-form} models.  Broadly
speaking, the parametric models use parameterized density
distributions to describe the arrangement of mass within the cluster,
typically associating massive dark matter halos with the positions of
bright cluster member galaxies. Therefore, parametric models rely (to
varying degrees) on the assumption that the cluster's dark matter can
be described by analytic forms such as NFW halos \citep{Navarro:1997},
or pseudo isothermal elliptical mass
distributions \citep[PIEMD][]{Kassiola:1993}.  Free-form models divide
the cluster into a grid, generally using a multi-scale grid to get
better sampling in denser regions. Each grid cell is assigned a mass
or a potential, and then the mass values are iteratively refined to
match the observed lensing constraints. In some cases an adaptive grid
is used so that the grid spacing itself can also be modified as the
model is iterated \citep[e.g.][]{Liesenborgs:2006,Merten:2009}.  
A more rigorous comparison of these
disparate lens modeling techniques is beyond the scope of this
work.  For a complete discussion of each model's
methodology, the reader is directed to the listed references.

In Figure~\ref{fig:LensingTest} the model predictions are plotted
alongside the observed magnification of SN \tomas, derived in
Section~\ref{sec:DistanceAndMagnification}.  To first order, this
comparison shows that these 15 models are largely consistent with each
other and with the observed magnification of SN \tomas. Naively
treating each model as an independent prediction for the magnification
(and ignoring their quoted uncertainties), one would get a mean for
the full sample of $\mu=2.5\pm0.4$.  This is separated from
the observed SN magnification by $\delta\mu/\mu=23\%$, which is
approximately a 1$\sigma$ difference.

In general, free-form methods tend to explore the model parameter
space more fully than parametric methods.  This is reflected in the
size of the error-bars in Figure~\ref{fig:LensingTest}.  The Williams
model (using GRALE) has the largest uncertainties, with a
68\% confidence range that encompasses the median magnification values
returned by all other models.  This is possibly the most complete
uncertainty estimate as it includes many systematic error sources
that are not incorporated into most other models. For example, this
implementation of GRALE accounts for mass sheet degeneracy by
introducing a free parameter in the reconstruction that represents an
arbitrarily scaled mass sheet, with normalization to be chosen by the
genetic algorithm.  More complex degeneracies are also accounted for
in this model through the use of basis
functions \citep{Liesenborgs:2006,Liesenborgs:2007,Mohammed:2014}.

@@

In spite of the great variety in input data and methodology, these
lens models overall are delivering consistent and fairly accurate
estimates of the magnification.  Seven of the tested models (CATS(v1),
GLAFIC, Zitrin-NFW, Merten, Bradac(v2), Lam(v2), Diego) report a
median value that falls within the 1$\sigma$ uncertainty range of the
measured magnification, and three of the free-form models (Williams,
Lam(v2), Diego) report uncertainties that overlap the central $\mu$ as
measured from both of the SN light curve fitters.  Furthermore, the
scatter amongst models provides a reasonable estimate of the
magnification uncertainty.

This is largely in agreement with the results
of \citetalias{Patel:2014} and \citet{Nordin:2014}.  The fact that the
model predictions are collectively within $1\sigma$ of the observed
magnification is especially encouraging for these Abell 2744
models. This is a merging cluster with a complex mass distribution,
and many of the models we are evaluating are preliminary products,
generated quickly and without access to the rich data from the
Frontier Fields program.

However, beyond this first-order agreement, there is a small
systematic bias apparent. All but two of the lens models
return median magnifications that are {\it higher} than the observed
value, and six of the models are discrepant by more than $1.5\sigma$.
It is important to emphasize that SN \tomas\ only samples a single
line of sight through the cluster, and this bias to higher
magnifications is minor.  Nevertheless, a systematic
shift of this nature is surprising, given the wide range of
modeling strategies, input data, and physical assumptions represented
by this set of models.  In the following subsections we examine
possible explanations for this small but nearly universal
bias.  We first consider whether a misinterpretation of the data on the SN
itself can account for the observed systematic bias, and then 
 examine the lens models.

@@

\subsection{Possible Errors in Lens Modeling}
\label{sec:LensModelComparison}

Having now ruled out misinterpetations of the SN data, we are left to
seek an explanation for the mild tension by scrutinizing the lens
models.  Here we evaluate four possible sources for a systematic bias to
appear in the lens models.

@@

\subsubsection{Misidentification of a multiple image}

\citet{Jauzac:2014c} proposed a correction
for the location of image 3.3, the third component of a triply-imaged
galaxy \citep{Merten:2011} at $z=3.98\pm0.02$ \citep{Johnson:2014}.
This possibly specious multiple image is only $\sim$10\arcsec\ from
the SN \tomas\ position, meaning a change in its location could have a
large impact on the predicted magnification for
SN \tomas\ \citep[see][for a quantitative discussion of this
effect]{Johnson:2014}.  However, \citet{Jauzac:2014c} also evaluated a
version of the CATS(v2) model in which this single source is left in
the original position and found that the location of image 3.3 does
not significantly affect the predicted magnifications.  Furthermore,
dividing our 15 models into those using the original location vs those
that adopt the new position, we find that both groups include models
that are within 1$\sigma$ of the observed $\mu$ as well as
significantly discrepant points.

\subsection{Quantity and Quality of Lensing Constraints}
\label{sec:LensModelComparison}

None of the possibilities described above can completely account for
the mild tension between the observed and predicted magnifications.
However, in lieu of a physical explanation, we can simply take a
pragmatic approach and seek to identify any shared characteristics of
the most accurate models.  In this way SN \tomas\ may serve as a guide
for optimizing future lens models for magnification predictions.

@@

As can be seen in Table~\ref{tab:LensModels} and
Figure~\ref{fig:LensingTest}, among the 7 models that yield
magnifications discrepant by more than 1$\sigma$ we find a variety of
methodologies (both parametric and free-form) and a mix of lensing
constraints (both strong-lensing only and strong+weak).  The left
panel of Figure~\ref{fig:Nspecz} shows that this subset of 7 also
includes a wide range in the number of multiply-imaged galaxies used
as strong-lensing constraints (from $\sim$10 to 50).  Given that the
discrepancy persists across all of these categories, none of them can
provide a satisfactory explanation for the small observed bias.

However, the right panel of Figure~\ref{fig:Nspecz} shows that 3 of
the most successful lens models (Bradac(v2), Lam(v2), Diego) do share
a common characteristic: they all use a set of multiply-imaged galaxies for
which a large fraction have spectroscopic redshifts. In practice,
this means that the strong-lensing constraints in these models are
defined more stringently. Other models use
a larger set of strong-lensing features, but the dearth of
spectroscopic constraints means that it is more likely that some of
those multiply-imaged galaxies are assigned with incorrect redshifts.

@@

It is relevant to note that the three models highlighted for their
accuracy here (Bradac(v2), Lam(v2), Diego) also happen to be the three
models that are not truly ``blind,'' in that the modelers were aware
of preliminary analysis of the SN magnification.  However, none of
these models used the SN magnification as an input constraint, none
were tuned to match the SN magnification, and the modelers did not
have a final measurement of the SN magnification at the time their
models were generated.

>>>>

In the case of lensed SNe, we can also reverse the experimental setup:
instead of using strong-lensing clusters to study distant SNe, we can
use the SNe as tools for examining the lenses \citep{Riehm:2011}.  The
most valuable transients for testing and improving cluster lens models
would be {\it strongly lensed} SNe that are resolved into multiple
images\ \citep{Holz:2001,Oguri:2003}.  Transients that are lensed into
multiple images can also become cosmological tools, as the measurement
of time delays between the images can provide cosmographic information
to constrain the Hubble parameter \citep{Refsdal:1964} and other
cosmological parameters \citep{Linder:2011}.  We have recently
observed the first example of a multiply-imaged SN \citep{Kelly:2015},
and expect to detect the reappearance of this object (called ``SN
Refsdal'') \change{within the next
year} \citep{Oguri:2015,Sharon:2015,Diego:2015}, delivering a precise
test of lens model predictions.  Although detections of such objects
are currently very unlikely \citep{Li:2012}, they will become much
more common in the next decade \citep{Coe:2009,Dobke:2009}, and may be
developed into an important new cosmological tool.

@@

One of the key values in observing standard candles behind
gravitational lenses is in addressing the problem of the
mass-sheet degeneracy \citep{Falco:1985,Schneider:1995}.  This
degeneracy arises because one can introduce into a lens model an
unassociated sheet of uniform mass in front of or behind the
lens, without disturbing the primary observable quantities.  For
example, take a lens model with a given surface mass density $\kappa$,
and then transform the surface mass density to
$\kappa^{\prime}=(1-\lambda)\kappa+\lambda$ for any arbitrary value
$\lambda$.  Both the $\kappa$ and $\kappa^{\prime}$ models will
produce exactly the same values for all positional and shear
constraints from strong and weak lensing \citep{Seitz:1997}.  When lensed background
sources are available across a wide range of redshifts (as is the case
for the Abell 2744 cluster discussed here), it should in principle be
possible to account for the simplest form of a single mass sheet
within that redshift range \citep{Bradac:2004}.  However, there are more complex versions
of positional constraint
degeneracies \citep{Schneider:2014,Liesenborgs:2012}.  Such
degeneracies do not extend to the {\it absolute} magnification of a
background source's flux and size.  Therefore, an absolute measurement
of magnification from a standard candle or a standard ruler
\citep{Sonnenfeld:2011}  can break these fundamental degeneracies
\citep{Holz:2001}.

@@

SN \tomas\ was discovered in Hubble Space Telescope (\HST) observations
with the Advanced Camera for Surveys (ACS) in the F606W and F814W
bands (V and i), collected on UT 2014 May 15 as part of the Hubble
Frontier Fields (HFF) survey (PI:J.Lotz,
HST-PID:13495).\footnote{\url{http://www.stsci.edu/hst/campaigns/frontier-fields}}
The HFF program is a 3-year Director's discretionary initiative that
is collecting 140 orbits of HST imaging (roughly 340 ksec) on six
massive galaxy clusters, plus 6 accompanying parallel fields.  Each
field is observed in 3 optical bands (ACS F435W, F606W and F814W) and
4 infrared (IR) bands (WFC3-IR F105W, F125W, F140W, and F160W),
although the optical and IR imaging campaigns are separated by $\sim$6
months. Abell 2744 was the first cluster observed, with IR imaging
spanning 2013 October-November, and optical imaging from 2014
May-July.  A composite image of the HFF data showing the SN is
presented in Figure~\ref{fig:DiscoveryImage}, and the
locations of the cluster center, the SN, and the presumed host galaxy
are given in Table~\ref{tab:Positions}.  The SN detection was made in
difference images constructed using template imaging of Abell 2744
from HST+ACS observations taken in 2009 (PI:Dupke, HST-PID:11689).

@@

From the preceding analysis, we conclude that \tomas\ is a Type Ia SN
at $z\approx1.35$.  At this redshift, the absorption at $\sim$8700\AA\
corresponds to the blended \ion{Ca}{2} H\&K \change{and \ion{Si}{2} $\lambda$3858} features.
\change{Generally referred to as the Ca H\&K feature, this} absorption is commonly seen in Type Ia SN spectra
near maximum light, although it is also prominent in the spectra of
Type Ib and Ic core collapse SNe (\CCSNe).  The red color of
the \tomas\ SED is qualitatively consistent with a redshift of $z>1$
-- although this information was not used by {\tt SNID} for the
template matching.  As we will see in
Section~\ref{sec:PhotometricClassification}, this spectral
classification of SN \tomas\ is reinforced by the photometric
information, which also supports classification as a Type Ia SN at
$z\approx1.35$.

@@

To refine the redshift and phase constraints on \tomas, we next fit
the spectrum with a custom spectral matching program that employs a
library of Type Ia SN SEDs.  This library is similar to the Type Ia
spectral set used by SNID, but also includes more recent SNe with
well-observed spectral time series that extend to rest-frame UV
wavelengths (e.g. SN 2011fe and 2014J).  We first use an approach
similar to the SNID algorithm: warping the pseudo-continuum of each
template spectrum by dividing out a 3$^{\rm rd}$-order polynomial to
match the observed SED of SN \tomas.  This approach will find
templates that have similar abundances and photospheric
velocities. With the redshift fixed at $z=1.3457$ we find the best fit
is a spectrum from the normal Type Ia SN 2014J \citep{Foley:2014},
with a $\chi^2$ per degree of freedom $\nu$ of $\chi^2/\nu=121.8/89$
shown in the top panel of Figure~\ref{fig:SpecFit}. \change{The excess
variance in this fit may be attributed to the intrinsic variation of
Type Ia SN spectra, which is more prominent at UV
wavelengths \citep[e.g.][]{Foley:2008b,Wang:2012} and is not fully
represented in the available template library.}  Allowing the redshift
as an additional free parameter, we still find results that are
consistent with the SNID fits and $<2\sigma$ from the host galaxy
redshift: $z=1.31\pm0.02$ and a phase of $0\pm3$ rest-frame days. The
best-fitting spectral template in this case is the normal Type Ia SN
2012cg \citep{Amanullah:2015} at $z=1.31$ with $\chi^2/\nu=98.5/88$
(second panel of Figure~\ref{fig:SpecFit}).

@@

With the type and redshift securely defined as a normal Type Ia SN at
$z=1.3457$, we now turn to fitting the light curve with Type Ia
templates to measure the distance
modulus \change{(\S\ref{sec:LightCurveFitting})} and then derive the
gravitational lensing
magnification \change{(\S\ref{sec:ControlSampleComparison})}.  In this
process, we would like to avoid introducing systematic uncertainties
inherent to any assumed cosmological model.  To that end, \change{in
section \ref{sec:ControlSampleComparison}} we
follow \citealt{Patel:2014} and define the magnification by comparing
the measured distance modulus of SN \tomas\ against an average
distance modulus derived from a ``control sample'' of unlensed Type Ia
SNe at similar redshift. This allows us to make only the minimal
assumption that the redshift-distance relationship for Type Ia SN is
smooth and approximately linear over a small redshift span, which
should be true for any plausible cosmological model.

@@

The unlensed sample comprises 22 spectroscopically confirmed Type Ia
SNe in the range $1.1<z<1.6$ from three HST surveys: 11 from the GOODS Higher-$z$ SN search\footnote{GOODS: the Great Observatories Origins Deep Survey, PI:Giavalisco, HST-PID:9425,9583} \citep{Strolger:2004,Riess:2007}, 7 from SCP\footnote{SCP: the Supernova Cosmology Project, PI:Perlmutter} \citep{Suzuki:2012}, and 4 from CANDELS\footnote{CANDELS: the Cosmic Assembly Near-infrared Deep Extragalactic Legacy Survey, PI:Faber \&\ Ferguson} \citep{Rodney:2012,Rodney:2014}.  Using the SALT2 and MLCS2k2
fitters as described above, we get distance modulus measures for every
object in this control sample.  We then fit a linear relationship for
distance modulus vs. redshift, and derive a prediction for the
distance modulus of a normal Type Ia SN at the redshift of SN \tomas\
(Figure~\ref{fig:MagnificationMeasurement}).
This predicted value is given in
Table~\ref{tab:MeasuredMagnification} under the ``Control'' column.
The difference between the observed distance of SN \tomas\ and this
control sample value is attributed to the magnification from
gravitational lensing:

@@

%\section{Discussion}
%\label{sec:Discussion}

\section{Comparison to Model Predictions}
\label{sec:Discussion}
\label{sec:ComparisonToModelPredictions}

@@

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{FIG/snTomas_magnifications}
\caption{ 
Comparison of the observed lensing magnification to predictions from
lens models. The vertical blue line shows the constraints from
SN \tomas\ derived in Section~\ref{sec:DistanceAndMagnification} using
the MLCS2k2 fitter, with a shaded region marking the total
uncertainty. Markers with horizontal error bars show the median
magnification and 68\%\ confidence region from each of the 15 lensing
models.  Circles indicate models that use only strong-lensing
constraints, while diamonds denote those that also incorporate
weak-lensing measurements.  \change{Models using a ``free-form''
approach are shown as open markers, while those in the ``parametric''
family are given filled markers.  The top half, with points in black, shows the nine
models that were constructed using only data available before the
start of the Frontier Fields observations. A black dashed line marks the
unweighted mean for these models, at $\mu=2.6$.
The lower six models in
green used additional input constraints, including new multiply imaged
systems and redshifts.  The final two points, with square orange
outlines, are the ``unblind'' models that were generated after the
magnification of the SN was known.  
The green dashed line marks the
unweighted mean for these six models, at $\mu=2.3$.
\label{fig:LensingTest} }}
\end{center}
\end{figure}

Table~\ref{tab:PredictedMagnifications} gives each model's predicted
magnification and uncertainty for a source at $z=1.3457$ and at the
position of \tomas.  These models represent a broad sampling of the
techniques and assumptions that can be applied to the modeling of mass
distributions in galaxy clusters. 
\change{We discuss some of the modeling choices here, but
a more rigorous comparison of these disparate lens modeling techniques
is beyond the scope of this work. For a complete discussion of each
model's methodology the reader is directed to the listed references.}

 Eight of these are so-called {\it parametric}\footnote{Although this
nomenclature is becoming standard in the literature, it is somewhat
misleading, as the pixels or grid cells in free-form models are
effectively parameters as well. Perhaps ``simply-parameterized'' would
be more accurate, though we adopt the more common usage here.}
models, and the remining seven are {\it free-form} models.  Broadly
speaking, the parametric models use parameterized density
distributions to describe the arrangement of mass within the cluster,
typically associating massive dark matter halos with the positions of
bright cluster member galaxies. Therefore, parametric models rely (to
varying degrees) on the assumption that the cluster's dark matter can
be described by analytic forms such as NFW halos \citep{Navarro:1997},
or pseudo isothermal elliptical mass
distributions \citep[PIEMD][]{Kassiola:1993}.

Free-form models divide
the cluster into a grid, generally using a multi-scale grid to get
better sampling in denser regions. Each grid cell is assigned a mass
or a potential, and then the mass values are iteratively refined to
match the observed lensing constraints. In some cases an adaptive grid
is used so that the grid spacing itself can also be modified as the
model is iterated \citep[e.g.][]{Liesenborgs:2006,Merten:2009}.  
\change{
In general, free-form methods tend to be more flexible than parametric methods,   
as reflected by the
size of the error-bars in Figure~\ref{fig:LensingTest}.
For example, the Williams
model (using GRALE) has the largest uncertainties, with a
68\% confidence range that encompasses the median magnification values
returned by all other models.  This is due in part to the 
 inclusion of many systematic error sources. For example, this
implementation of GRALE accounts for mass sheet degeneracy by
introducing a free parameter in the reconstruction that represents an
arbitrarily scaled mass sheet. More complex degeneracies are also accounted for
in this model through the use of basis
functions \citep{Liesenborgs:2006,Liesenborgs:2007,Mohammed:2014}. 
}

@@

In Figure~\ref{fig:LensingTest} the model predictions are plotted
alongside the observed magnification of SN \tomas, derived in
Section~\ref{sec:DistanceAndMagnification}.  To first order, this
comparison shows that these 15 models are largely consistent with each
other and with the observed magnification of SN \tomas. \change{If we
take the first set of 9 pre-HFF models and naively treat each as an
independent prediction for the magnification (they are not all
independent) and if we ignore their quoted uncertainties, we would get
an unweighted mean for the pre-HFF sample of $\mu_{\rm
pre}=2.6\pm0.4$.  This is separated from the observed SN magnification
by $\delta\mu/\mu=28\%$, which is approximately a 1.5$\sigma$
difference. The addition of deeper imaging and new spectroscopic
constraints from the HFF program and affiliated efforts should be
expected to improve the lens models, and in fact we see that it does:
the mean for the post-HFF sample is $\mu_{\rm post}=2.3\pm0.6$, fully
consistent with the observed SN magnification.}

This is largely in agreement with the results
of \citetalias{Patel:2014} and \citet{Nordin:2014}, \change{where
model predictions were found to be in reasonably good agreement with a
set of 3 lensed SNe from the CLASH program.  The general agreement
between the model predictions and the SN measurement} is especially
encouraging for these Abell 2744 models. This is a merging cluster
with a complex mass distribution, and \change{the SN is located
outside of the strong-lensing region where the models are most tightly
constrained.}

%In spite of the great variety in input data and methodology, these
%lens models overall are delivering consistent and fairly accurate
%estimates of the magnification.  Seven of the tested models (CATS(v1),
%GLAFIC, Merten, Zitrin-NFW, Bradac(v2), Diego, Lam(v2)) report a
%median value that falls within the 1$\sigma$ uncertainty range of the
%measured magnification, and three of the models (Williams,
%Diego, Lam(v2)) report uncertainties that overlap the central $\mu$ as
%measured from both of the SN light curve fitters.

However, beyond this first-order agreement, there is a small
systematic bias apparent. All but two of the lens models return median
magnifications that are {\it higher} than the observed value, and six
of the models are discrepant by more than $1.5\sigma$. \change{These
six discrepant models are all biased to higher magnifications. They
are found in both the pre-HFF and post-HFF models, in the parametric
and free-form families, and among the strong-lensing-only and the
strong+weak subsets.}  It is important to emphasize that SN \tomas\
only samples a single line of sight through the cluster, and this bias
to higher magnifications is minor.  Nevertheless, a systematic shift
of this nature is surprising, given the wide range of modeling
strategies, input data, and physical assumptions represented by this
set of models.  In the following subsections we examine possible
explanations for this small but nearly universal bias.  We first
consider whether a misinterpretation of the data on the SN itself can
account for the observed systematic bias, and then examine the lens
models.

@@

\subsection{Possible Errors in Lens Model Inputs}
\label{sec:LensModelInputs}

Having \change{found no evidence for} misinterpetations of the SN
data, we are left to seek an explanation for the mild tension by
scrutinizing the lens models.  Here we evaluate four
possible \change{ways in which the input data or assumptions of these
lens models could lead to a systematic bias in the magnification.}

@@

\subsubsection{Misidentification of a multiple image}
\citet{Jauzac:2014c} proposed a correction
for the location of image 3.3, the third component of a triply-imaged
galaxy \citep{Merten:2011} at $z=3.98\pm0.02$ \citep{Johnson:2014}.
This possibly specious multiple image is only $\sim$10\arcsec\ from
the SN \tomas\ position, meaning a change in its location could have a
large impact on the predicted magnification for
SN \tomas\ \citep[see][for a quantitative discussion of this
effect]{Johnson:2014}.  However, \citet{Jauzac:2014c} also evaluated a
version of the CATS(v2) model in which this single source is left in
the original position and found that the location of image 3.3 does
not significantly affect the predicted magnifications.  Furthermore,
dividing our 15 models into those using the original location vs those
that adopt the new position, we find that both groups include models
that are within 1$\sigma$ of the observed $\mu$ as well as
significantly discrepant points.

\change{
\subsection{Best Practices for Magnification Prediction}
\label{sec:BestPractices}
}

\change{
None of the possible errors in lens model inputs described above can
completely account for the mild tension between the observed and
predicted magnifications.  However, we do have several models that
agree with the measured SN magnification.  In this section we take a
simple pragmatic approach and seek to identify any shared
characteristics of the most accurate models.  In this way SN \tomas\
may serve as a guide toward optimizing future lens models for
magnification predictions.  }

@@

\change{
\subsubsection{Quantity and Quality of Strong Lensing Constraints}
\label{sec:QuantityAndQuality}
}

\change{
Increasing the number of multiply imaged galaxies used as
strong-lensing constraints should in principle lead to improvements in
the accuracy and precision of magnification predictions.  We have
already seen some suggestion of this from
Figure~\ref{fig:LensingTest}, which shows that the post-HFF models are
collectively better than the pre-HFF models at predicting the
SN \tomas\ magnification.  
The left panel of Figure~\ref{fig:Nspecz} isolates this variable by
plotting the tension between each model prediction and the measured SN
magnification against the number of multiply-imaged galaxies used as
strong-lensing constraints. There is no clear correlation here, and in
fact the most significant outlier (CATS(v2)) is actually the one using
the most multiply-imaged systems.  
}

\change{
In the right panel of
Figure~\ref{fig:Nspecz} we examine an alternative explanation by plotting 
the model tension against the fraction of strong-lensing constraints that have a spectroscopic redshift. 
Here we see a suggestion of a trend, and find that two of the most successful lens
models (Bradac(v2) and Lam(v2)) do share a common characteristic: they
both use a set of multiply-imaged galaxies for which a large fraction
have spectroscopic redshifts. In practice, this means that the
strong-lensing constraints in these models are defined more
stringently. Other models use a larger set of strong-lensing features,
but the dearth of spectroscopic constraints means that it is more
likely that some of those multiply-imaged galaxies are assigned with
incorrect redshifts.
}

@@

It is relevant to note that the two models highlighted for their
accuracy here (Bradac(v2) and Lam(v2)) also happen to be the two
models that are not truly ``blind,'' in that the modelers were aware
of preliminary analysis of the SN magnification.  However, neither of
these models used the SN magnification as an input constraint, neither
were consciously tuned to match the SN magnification, and the modelers did not
have a final measurement of the SN magnification at the time their
models were generated.

%\subsubsection{Source Plane Minimization}
%
%A third model that stands out for delivering both an accurate and fairly precise estimate of the magnification 
%
%A general tendency towards high model magnification can result when
%lens models are minimized in the source plane (rather than the image
%plane).  As the overall magnification of a given model
%increases, differences in angle between the delensed
%images of the same source become smaller, corresponding to an overly shallow lensing mass profile. In
%the limit of only a small number of multiple images that cover a
%limited range of source redshift, this results in a tendency to force all
%sources toward the same location in the source plane, thereby
%maximizing the magnification.
%
%Minimizing in the image plane avoids this well known bias but is much
%more difficult (Broadhurst etal 2005 ApJ 621,53, Zitrin etal 2009,
%MNRAS 396,1985) as model generated images must be searched for in the
%image plane near the corresponding observed images.  
%
%Alternatively, one may guard against this bias using a modified source plane minimization that is
%constrained not to generate overly small re-lensed sources  (Sendra etal 2012,MNRAS 437,2642,Diego etal
%2014).
%
%For models that incorporate HFF
%data, this bias should be reduced, as many
%multiply images spread over a wide redshift range will lessen the bias
%towards high magnification. 
%
%
%
%\subsubsection{Model Flexibility}
%
%Let us first define a successful magnification prediction as one whose
%quoted uncertainties overlap the measured $\mu=2.03$ from the SN.
%There are three models that meet this criteria: Williams, Diego, and
%Lam(v2). All of these are free-form models
